---
title: "Forecasting Weekly Sales for Walmart"
author: "Daniel Garcia, Nicole Pham"
date: "2022-12-05"
output: html_document
---

```{r setup, include=FALSE, error=TRUE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
```


#####################################################################################
#######----------- Libraries/data importing/cleaning/exploring ----------------######
#####################################################################################



### Libraries
```{r}
# importing libraries
library(tidyverse)
library(skimr)
library(lubridate)
library(MASS)
library(rcompanion)
library(tidymodels)
```

### Data importing/cleaning/exploring
```{r}
# importing data set
Walmart= read_csv("C:\\Users\\dgray\\OneDrive\\Stats\\Stat 471 R programming\\R projects\\Walmart.csv")

#Walmart= read_csv("/Users/nicolepham/Documents/STAT/STAT 471/Walmart.csv", show_col_types = FALSE)
glimpse(Walmart)
```


```{r}
# changing Date variable to date data structure
# format= year/Month/Day

Walmart<-Walmart %>%
  mutate(Date=dmy(Date))
glimpse(Walmart)
```
```{r}
range(Walmart$Date)
```


```{r}
Walmart %>% 
  dplyr::select(Weekly_Sales, Temperature, Fuel_Price, CPI, Unemployment) %>% 
  skim()
```



###################################################################################
#----------------------------------- EDA -----------------------------------------#
###################################################################################


#### 1) Weekly Sales

```{r}
ggplot(Walmart, aes(x=Weekly_Sales)) + 
  geom_histogram(bins=24, fill="#CC9900", color="black") +
  labs(title="Distribution of Weekly Sales", x="Weekly Sales")
```

```{r}
shapiro.test(head(Walmart$Weekly_Sales, 500))
```


Distribution is not normally distributed by Shapiro Test.

```{r}
mean(Walmart$Weekly_Sales)
```

```{r}
median(Walmart$Weekly_Sales)
```



Many outliers, but mean/median is around $1,000,000 in weekly Sales. Median is a better average here, given the outliers.

Mean: $1,046,965
Median: $960,746




#### 2) Weeks with Holidays

```{r}
# using filter function to see which weeks are holiday weeks
holiday_flag = Walmart %>% 
  dplyr::filter(Holiday_Flag==1)
n_unique(holiday_flag$Date)
```

```{r}
holiday_flag
```




```{r}
ggplot(Walmart, mapping=aes(x=Weekly_Sales, color=factor(Holiday_Flag))) +
  geom_histogram(position="dodge", bins=14) +
  labs(title="Histogram of weekly sales by Holiday weeks", x="Weekly Sales", color="Holiday flag")
```


```{r}
ggplot(Walmart, aes(x=as_factor(Holiday_Flag), y=Weekly_Sales, color=factor(Holiday_Flag)))+
  geom_boxplot() +
  labs(title = 'BoxPlot of Weekly Sales by Holiday Weeks', y='Weekly sales', x='Holiday Flag', color="Holiday Flag")
```

slight increase in median sales for holiday weeks. Outliers in non-holiday weeks. 


```{r}
ggplot(data=holiday_flag, mapping=aes(x=Temperature,y=Weekly_Sales, color=factor(Date))) +
  geom_point() +
  geom_smooth(method=lm, formula=y~x, se=FALSE) + 
  labs(title="Scatterplot of Weekly Sales vs Temperature by Holiday weeks", y="Weekly Sales", color="Date")
```

```{r}
ggplot(data=holiday_flag, mapping=aes(x=Fuel_Price,y=Weekly_Sales, color=factor(Date))) +
  geom_point() +
  geom_smooth(method=lm, formula=y~x, se=F) + 
  labs(title="Scatterplot of Weekly Sales vs Fuel Price by Holiday weeks", y="Weekly Sales", color="Date")
```
#### 3) By Store

```{r}
# checking number of stores
unique(Walmart$Store)
```

```{r}
# setting store number as factor
Walmart$Store = as.factor(Walmart$Store) 
class(Walmart$Store)
```

```{r}
# boxplot of weekly sales by Store
  ggplot(Walmart, aes(Weekly_Sales, reorder(Store,
                                   FUN = median, Weekly_Sales))) + 
  geom_boxplot(color="Dark green") + 
  labs(title="Boxplot of Weekly sales by Store", x="Weekly Sales", y="Store number")
```
Store 4 has the highest median weekly sales. store 33 has the lowest. Store 28 looks like  it has the highest variability, given the extent of the box length. 

#### 4) Weekly Sales By Consumer Price Index

```{r}
ggplot(Walmart, aes(x=CPI, y=Weekly_Sales)) + 
  geom_point(alpha=.1) +
  labs(title="Scatterplot of Weekly Sales vs Consumer Price Index", y="Weekly Sales")
```
appears that CPI doesnt effect weekly sales, since weekly sales seem equally distributed across all CPI ranges

```{r}
ggplot(Walmart, aes(x=CPI)) +
  geom_histogram(bins=29, fill="orange", color="black") +
  labs(title = 'Histogram of CPI', y='Count', x='CPI') 
```

most populated CPI ranges: 125-150, 180-225

#### 5) Unemployment

```{r}
ggplot(Walmart, mapping=aes(x=Unemployment)) +
  geom_histogram(bins=12, fill="dark green", color="white") + 
  labs(title="Histogram of Unemployment rate", )
```


```{r}
ggplot(Walmart, mapping=aes(x=Unemployment, y=Weekly_Sales, reorder(Store,
                                   FUN = median, Weekly_Sales), color=Store)) + 
  geom_point(alpha=.15) +
  geom_smooth(method=lm, formula= y~x, color="grey") +
  labs(title="Scatterplot of Weekly Sales vs. Unemployment rate", x="Unemployment rate", y="Weekly Sales", color="Store number")
```




Seems like a moderate decrease in weekly sales as the unemployment rate increases. 



#### Time

```{r}
# grouping by date and finding mean of all 45 walmart stores per week// also outputting if the week is a holiday week
Mean_wkly_sales= Walmart %>% 
  group_by(Date, Holiday_Flag) %>% 
  summarise(Weekly_Sales = mean(Weekly_Sales, na.rm=TRUE))
Mean_wkly_sales
```

```{r}
# line graph of all average weekly sales, highlighting those that surpassed $1.2M
ggplot(Mean_wkly_sales, mapping=aes(Date, Weekly_Sales)) + 
  geom_point(aes(color=Weekly_Sales>1200000), show.legend=T) +
  geom_line(group=2, color="grey") +
  labs(title="Weekly Sales by Date", subtitle="Showing average weekly sales greater than 1.3M", y="weekly sales", color="Weeks with Weekly Sales > 1.2M") +
  theme_bw() 
```

Certain Dates have outliers. I've defined outliers to be weekly sales that surpass $1.4M Let's look further into this and see if they are holiday days.

```{r}
# tibble showing dates that mean weekly sales surpass $1.2M
all_extremes = Mean_wkly_sales %>% 
  filter(Weekly_Sales>1200000) 
all_extremes

# counting number of times mean weekly sales passed $1.2M
nrow(all_extremes)
```



```{r}
# tibble showing holiday weeks that passed $1.2M in average weekly sales
num_holiday_extremes = all_extremes %>% 
  filter(Holiday_Flag==1) 
nrow(num_holiday_extremes)
num_holiday_extremes
```


```{r}
# proportion of times weekly sales surpassed $1.2M in a holiday week 
p= nrow(num_holiday_extremes)/nrow(all_extremes)
p
```

25% of the days that had weekly sales greater that $1.2M were flagged as Holiday weeks. 



```{r}
# Line graph showing average weekly sales by date with holiday flags highlighted

ggplot(Mean_wkly_sales, mapping=aes(Date, Weekly_Sales)) + 
  geom_point(aes(color=factor(Holiday_Flag)), show.legend=T) +
  geom_line(group=2, color="grey") +
  labs(title="Weekly Sales in the Extreme", subtitle="Showing Holiday flags for average weekly sales greater than 1.3M", y="weekly sales", color='Holiday Flag') +
  theme_bw()
```


#### Correlation Coefficients for all numeric variables

```{r}
cor(Walmart[, c("Weekly_Sales", "Temperature", "CPI", "Fuel_Price", "Unemployment")])
```


####################################################################################################
#----------------------------------- Regression Section -------------------------------------------#
####################################################################################################


## Regression on average weekly sales across all Walmarts
#### General Linear Regression
```{r}
# fitting a general linear regression model

summary(fitted.model<-glm(Weekly_Sales ~ Store + Holiday_Flag + Temperature + Fuel_Price + CPI + Unemployment, data=Walmart, family=gaussian(link=identity)))
```

not a good model since the response is not normally distributed. We will use a box-cox power transformation






#### Box-Cox Power Transformation
```{r}
mean_all = Walmart %>% 
  group_by(Date) %>% 
  summarize(Weekly_Sales = mean(Weekly_Sales),
            Temperature= mean(Temperature),
            CPI= mean(CPI),
            Fuel_Price= mean(Fuel_Price),
            Unemployment=mean(Unemployment))
mean_all
```
```{r}
shapiro.test(head(Walmart$Weekly_Sales, 400))
```

```{r}
# fitting box-cox transformation
boxcox_fit<-boxcox(Weekly_Sales ~ Temperature + CPI + Fuel_Price + Unemployment,
                             data=Walmart, lambda=seq(-2,2,1/4), interp=FALSE)
```


```{r}
boxcox.data = data.frame(boxcox_fit$x, boxcox_fit$y)
```


```{r}
ordered.data<- boxcox.data[with(boxcox.data, order(-boxcox_fit.y)),]
ordered.data[1,]

```

```{r}

wkly_sales= Walmart$Weekly_Sales
tr_wkly_sales = log(Walmart$Weekly_Sales)
head(tr_wkly_sales)
```

```{r}
plotNormalHistogram(tr_wkly_sales)
```

```{r}
shapiro.test(head(tr_wkly_sales, 450))
```
Box-Cox Power transformation is not a good fit. Trying Gamma model next for right-skewed distributions. 

#### Gamma Regression

```{r}
# setting reference level for holiday_flag and Store
Store_ = relevel(factor(Walmart$Store), ref=5)
Holiday_Flag_ = relevel(factor(Walmart$Holiday_Flag), ref="0")
```


```{r}
# fitting Gamma regression model
summary(fitted.model<-glm(Weekly_Sales ~ Store_ + Holiday_Flag_ + Temperature + Fuel_Price + CPI + Unemployment, data=Walmart, family=Gamma(link=log)))
```






```{r}
# fitting null model to perform deviance test 
null.model<- glm(Weekly_Sales ~ 1, data=Walmart, family=Gamma(link=log))


# deviance test
print(deviance<- -2*(logLik(null.model)-logLik(fitted.model)))

```


```{r}
# p value of deviance test
# if p value> .05, null model is better fit. If p value< .05, fitted model is better fit.
p.value<-pchisq(deviance, df=50, lower.tail=FALSE)
p.value
```



```{r}
sample_n(Walmart, 1)
```



```{r}
print(predict(fitted.model, data.frame(Store_="20", Holiday_Flag_="0", Temperature=50.52, Fuel_Price=2.917, CPI=205.0628, Unemployment=7.484), type="response"))

```
```{r}
sample_n(Walmart, 1)
```

```{r}
print(predict(fitted.model, data.frame(Store_="1", Holiday_Flag_="0", Temperature=86.83, Fuel_Price=3.682, CPI=215.4834, Unemployment=7.962), type="response"))
```






###################################################################################
################################ Analysis #########################################
###################################################################################


```{r}
# Changing Holiday flag to 1 during holiday times where holiday weeks are considered 0
x = Walmart %>% 
  filter((Date=="2010-12-10" | Date=="2010-12-17" | Date=="2010-12-24" | Date=="2011-12-09" | Date=="2011-12-16" | Date=="2011-12-23") & Holiday_Flag==0) %>% 
  mutate(Holiday_Flag = replace(Holiday_Flag, Holiday_Flag==0, 1))
x
```


```{r}
# modifying Walmart data to delete rows with Holiday weeks that are not flagged
rem_hol = subset(Walmart, (Date!="2010-12-10" & Date!="2010-12-17" & Date!="2010-12-24" & Date!="2011-12-09" &
                     Date!="2011-12-16" & Date!="2011-12-23"))
rem_hol %>% 
  filter(Date=="2011-12-09")
```
```{r}
# appending new holiday weeks to modified Walmart data
Walmart_new= rbind(x, rem_hol)
Walmart_new %>% 
  filter(Date=="2010-12-10")
```

```{r}
# calculating mean of each sales week using new Holiday flags
Walmart_new_mean= Walmart_new %>% 
  group_by(Date, Holiday_Flag) %>% 
  summarise(Weekly_Sales = mean(Weekly_Sales, na.rm=TRUE))
Walmart_new_mean
```



```{r}
# Line graph showing average weekly sales by date with new holiday flags highlighted

ggplot(Walmart_new_mean, mapping=aes(Date, Weekly_Sales)) + 
  geom_point(aes(color=factor(Holiday_Flag)), show.legend=T) +
  geom_line(group=2, color="grey") +
  labs(title="Weekly Sales in the Extreme", subtitle="Showing Holiday flags for average weekly sales greater than 1.2M", y="weekly sales", color='Holiday Flag') +
  theme_bw()
```

We now see that all extremes over 1.2M are during the holiday times. Specifically, from thanksgiving week to Christmas.  

#### Separating outlier weeks

```{r}
Walmart_year = Walmart |>
  separate(Date, c("Year", "Month", "Day"), sep = "-", remove = FALSE) |>
  transform(Year = as.factor(Year), Month = as.factor(Month), Day = as.factor(Day))

ggplot(Walmart_year, aes(x=Year, y=Weekly_Sales))+
  geom_boxplot() +
  labs(title = 'BoxPlot of Weekly Sales by Year', y='Weekly sales', x='Year')

#finding and saving the highest weekly sale in 2012 to differentiate outliers in 2010 and 2011
Walmart_year2012 = Walmart_year |> 
  group_by(Year, Weekly_Sales) |> 
  filter(Year == 2012) 
max2012 = max(Walmart_year2012$Weekly_Sales)

new_extremes = Walmart_year |> 
  filter(Weekly_Sales > max2012) 

#most of the high sales are from thanksgiving weeks to xmas weeks
ggplot(data=new_extremes, mapping=aes(x=Date,y=Weekly_Sales, color=factor(Year))) +
  geom_point() +
  labs(title="Scatterplot of Outliers of Weekly Sales vs Date", y="Weekly Sales")
```

### Separating the date
##### BoxPlot of Weekly Sales by Year

```{r}
#looking at data by year
Walmart_year = Walmart |>
  separate(Date, c("Year", "Month", "Day"), sep = "-", remove = FALSE) |>
  transform(Year = as.factor(Year), Month = as.factor(Month), Day = as.factor(Day))

ggplot(Walmart_year, aes(x=Year, y=Weekly_Sales))+
  geom_boxplot() +
  labs(title = 'BoxPlot of Weekly Sales by Year', y='Weekly sales', x='Year')
```
### Bootstrapping
```{r}
#grouping by month to get the mean sales by month
Walmart_month = Walmart_year |>
  dplyr:::select(Month, Weekly_Sales) |>
  group_by(Month) |>
  summarize(mean_sales = mean(Weekly_Sales))

#bootstrapping sales each month with 50 reps
virtual_resampled_means <- Walmart_month |>
  rep_sample_n(size = 12, replace = TRUE, reps = 50) |>
  group_by(replicate) |>
  summarize(mean = mean(mean_sales))

ggplot(virtual_resampled_means, aes(x = mean)) +
  geom_histogram(binwidth = 10000, color = "white", boundary = 1300000) +
  labs(x = "sample mean sales",
       title = "Bootstrap distribution of the sample mean")
```

```{r}
#bootstrapping sales each month with 100 reps
virtual_resampled_means <- Walmart_month |>
  rep_sample_n(size = 12, replace = TRUE, reps = 100) |>
  group_by(replicate) |>
  summarize(mean = mean(mean_sales))

ggplot(virtual_resampled_means, aes(x = mean)) +
  geom_histogram(binwidth = 10000, color = "white", boundary = 1300000) +
  labs(x = "sample mean sales",
       title = "Bootstrap distribution of the sample mean")
```

```{r}
#bootstrapping with 1000 reps
virtual_resampled_means3 <- Walmart_month |>
  rep_sample_n(size = 12, replace = TRUE, reps = 1000) |>
  group_by(replicate) |>
  summarize(mean = mean(mean_sales))

ggplot(virtual_resampled_means3, aes(x = mean)) +
  geom_histogram(binwidth = 10000, color = "white", boundary = 1300000) +
  labs(x = "sample mean sales",
       title = "Bootstrap distribution of the sample mean")
```

From above:
Mean: $1,046,965
Median: $960,746

The mean is very close to the mean of the bootstrap distribution with 1000 reps. 

```{r}
#confidence interval using the percentile method
quantile(virtual_resampled_means3$mean, c(0.025, 0.975))

#visual of the confidence interval
ggplot(virtual_resampled_means3, aes(x = mean)) +
  geom_histogram(binwidth = 10000, color = "white", boundary = 1300000) +
  geom_vline(xintercept = c(1005037, 1098265), color = "blue") +
  labs(x = "sample mean sales",
       title = "Bootstrap distribution of the sample mean",
       subtitle = "Percentile method confidence interval")
```

Confidence Interval(Percentile Method): ($1,006,330, $1,102,230)

```{r}
#confidence interval using the standard error method
#middle 95% of values of bootstrap distribution
virtual_resampled_means3 |>
  summarize(SE = sd(mean))

#visual of the confidence interval
ggplot(virtual_resampled_means3, aes(x = mean)) +
  geom_histogram(binwidth = 10000, color = "white", boundary = 1300000) +
  geom_vline(xintercept = c(998285.97, 1095644.03), color = "blue") +
  labs(x = "sample mean sales",
       title = "Bootstrap distribution of the sample mean",
       subtitle = "Standard error confidence interval")
```

Confidence Interval(Standard Error Method): ($998,285.97, $1,095,644.03)

```{r}
#visual of both confidence intervals
ggplot(virtual_resampled_means3, aes(x = mean)) +
  geom_histogram(binwidth = 10000, color = "white", boundary = 1300000) +
  geom_vline(xintercept = c(1005037, 1098265), color = "blue") +
  geom_vline(xintercept = c(998285.97, 1095644.03), color = "red") +
  labs(x = "sample mean sales",
       title = "Bootstrap distribution of the sample mean",
       subtitle = "blue = percentile method, red = standard error method")
```

############# other data explorations ######################

```{r}
ggplot(data = Walmart_year, mapping = aes(x = Temperature, y = Weekly_Sales)) +
  geom_point() +
  geom_smooth(method = lm, formula = y ~ x, se = FALSE) +
  facet_wrap(~ Year, ncol = 1) +
  labs(title="Scatterplot of Weekly Sales vs Temperature by Year", y = "Weekly Sales") 
```

```{r}
ggplot(data = Walmart_year, mapping = aes(x = Fuel_Price, y = Weekly_Sales)) +
  geom_point() +
  geom_smooth(method = lm, formula = y ~ x, se = F) + 
  facet_wrap(~ Year, ncol = 1) +
  labs(title="Scatterplot of Weekly Sales vs Fuel Price by Year", x = "Fuel Price", y = "Weekly Sales")
```




